---
title: "Final LTDS Report"
date: "`r Sys.Date()`"
output: pdf_document
params:
  effort: "NA"
  nburr: "NA"
  burr.w.mn: "NA"
  burr.w.min: "NA"
  burr.w.max: "NA"
  comm.comm: "NA"
  occ: "NA"
  unocc: "NA"
  unk: "NA"
  site.name: "NA"
  tort.dens: "NA"
  tort.pop: "NA"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}

library(raster)
library(rgeos)
library(rgdal)
library(dplyr)
library(readr)
library(nimble)
library(terra)
library(ggplot2)
library(gt)
```
# LTDS Summary - `r params$site.name`

Final report for line transect distance sampling survey with burrow scoping.

We walked a total of __`r format(params$effort, scientific=FALSE)`__ m during our full LTDS survey, where we found a total of __`r params$nburr`__ burrows. We calculated a density of __`r params$tort.dens`__ tortoises/acre, resulting in approximately __`r params$tort.pop`__ tortoise at the recipient site. We calculated a mean burrow width of __`r params$burr.w.mn`__ mm (range = __`r params$burr.w.min`__ - __`r params$burr.w.max`__ mm).

```{r burr.w.hist, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
  ltds_data <- read_csv("./LTDS_example_data.csv")
  #Create frequency plot of burrow widths
  wh.plot <- ggplot(data = ltds_data, aes(x = Burrow_width)) +
    #geom_density(fill = "grey") +
    geom_histogram(aes(fill = ..density..), color = "black",
                   binwidth = 20, alpha = 0.8) +
    scale_fill_viridis_c(option = "viridis", begin = 0.3, end = 0.6) +
    ylab("Frequency") +
    xlab("Burrow width") +
    ggtitle("Distribution of burrow widths") +
    theme_classic() +
    theme(legend.position = "none")

  wh.plot
```

_Figure 1_. Histogram representing the distribution of burrow width measurements from detected burrows during the full LTDS survey. Data includes measurements from all burrow conditions and occupancy status. 


```{r burr.w.dens, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
  #Create density plot of burrow widths
  wd.plot <- ggplot(data = ltds_data, aes(x = Burrow_width)) +
    geom_density(aes(y = ..density..), fill = "grey") +
    #geom_histogram(aes(fill = ..density..), color = "black",
    #               binwidth = 20, alpha = 0.8) +
    scale_fill_viridis_c(option = "viridis", begin = 0.3, end = 0.6) +
    ylab("Density") +
    xlab("Burrow width") +
    ggtitle("Distribution of burrow widths") +
    theme_classic() +
    theme(legend.position = "none")

  wd.plot
```

_Figure 2_. Density plot representing the distribution of burrow width measurements from detected burrows during the full LTDS survey. Data includes measurements from all burrow conditions and occupancy status. 

We found __`r params$occ`__ __occupied__ burrows, __`r params$unocc`__ __unoccupied__ burrows, and __`r params$unk`__ __unknowns__.

```{r occu.figure, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
#Plotting burrow types
o.plot <- ggplot(data = ltds_data, aes(x = Occupied)) +
  #geom_density(aes(y = ..density..), fill = "grey") +
  geom_histogram(aes(fill = Occupied), color = "black",
                 alpha = 0.8, stat = "count") +
  scale_fill_viridis_d(option = "viridis", begin = 0.3, end = 0.6) +
  scale_x_discrete(limits = c("No","Yes","Unknown")) +
  ylab("Frequency") +
  xlab("Burrow occupied?") +
  ggtitle("Burrow occupancy status") +
  theme_classic() +
  theme(legend.position = "none")

o.plot
```

_Figure 3_. Histogram showing the occupancy status from detected burrows during the full LTDS survey. 

The most common commensal species discovered in burrows during scoping was the __`r params$comm.comm`__. A breakdown of the commensal species detection can be seen in the following table and figure:

_Table 1_. Name and total number of commensal species found inside gopher tortoise burrows during full LTDS survey.
```{r commensals.table, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
#List of commensals
comm.sp <- unique(ltds_data$Commensals)
comm.sp <- na.omit(comm.sp)

#Remove NA's from ltds data
ltds.na <- ltds_data %>%
  na.omit()

#Create numbering dataframe for reordering ggplot
#Make list for new val column
com.list <- vector(mode = "list", length = length(comm.sp))
for (i in 1:length(comm.sp)) {
  new.sum <- sum(ltds.na$Commensals == comm.sp[i])
  com.list[i] <- new.sum
}
#Turn list into string
val.data <- do.call(rbind.data.frame, com.list)
colnames(val.data) <- c("val")
#Join new variables
comm.data <- cbind(comm.sp, val.data)
colnames(comm.data) <- c("Species", "No.found")
new.data <- comm.data[order(comm.data$No.found),]

#Create table
comm.data.gt <- new.data %>%
  gt() %>%
  tab_header("Commensal species discovered")

comm.data.gt
```


```{r commensals.fig, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
#List of commensals
comm.sp <- unique(ltds_data$Commensals)
comm.sp <- na.omit(comm.sp)

#Remove NA's from ltds data
ltds.na <- ltds_data %>%
  na.omit()

#Create numbering dataframe for reordering ggplot
#Make list for new val column
com.list <- vector(mode = "list", length = length(comm.sp))
for (i in 1:length(comm.sp)) {
  new.sum <- sum(ltds.na$Commensals == comm.sp[i])
  com.list[i] <- new.sum
}
#Turn list into string
val.data <- do.call(rbind.data.frame, com.list)
colnames(val.data) <- c("val")
#Join new variables
comm.data <- cbind(comm.sp, val.data)

#Plotting number of commensals
c.plot <- ggplot(data = comm.data,) +
  geom_col(aes(x = reorder(comm.sp, +val), y = val, fill = val), color = "black") +
  scale_fill_viridis_c(option = "viridis", begin = 0.3, end = 0.6) +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  ylab("Frequency") +
  xlab("Commensal species") +
  ggtitle("Number of commensal species discovered") +
  theme_classic() +
  theme(legend.position = "none")

c.plot
```

_Figure 4_. Histogram showing the unmber of commensal species found within detected burrows during the full LTDS survey. 

```{r det.dist.hist, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
#Create detection distance plot
dh.plot <- ggplot(data = ltds_data, aes(x = Distance)) +
  #geom_density(fill = "grey") +
  geom_histogram(aes(fill = ..density..), color = "black",
                 binwidth = 1, alpha = 0.8) +
  scale_fill_viridis_c(option = "viridis", begin = 0.3, end = 0.6) +
  ylab("Frequency") +
  xlab("Distance") +
  ggtitle("Detection distances") +
  theme_classic() +
  theme(legend.position = "none")

dh.plot
```

_Figure 5_. Histogram showing the perpendicular distance from transect lines of all detected burrows during the full LTDS survey. 

```{r det.dist.dens, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
#Create detection distance plot
dd.plot <- ggplot(data = ltds_data, aes(x = Distance)) +
  geom_density(aes(y = ..density..), fill = "grey") +
  #geom_histogram(aes(fill = ..density..), color = "black",
  #               binwidth = 20, alpha = 0.8) +
  scale_fill_viridis_c(option = "viridis", begin = 0.3, end = 0.6) +
  ylab("Density") +
  xlab("Detection distance") +
  ggtitle("Detection distances") +
  theme_classic() +
  theme(legend.position = "none")

dd.plot
```

_Figure 6_. Density plot showing the perpendicular distance from transect lines of all detected burrows during the full LTDS survey. 

End of report. 
