---
title: "LTDSMethod vignette - Max Dolton Jones, 2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LTDSMethod vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Introduction to LTDSMethod

The LTDSMethod package was designed for several main tasks:

* Designing line transect distance sampling pilot surveys
* Using pilot survey methodology to inform full LTDS surveys
* Estimate the density of a species within a study area
* Creating a summary of survey effort and results

More specifically, the package was initially intended as a tool to aid monitoring
efforts for translocated gopher tortoises throughout Florida. As a result, some
of the functions are highly specific to gopher tortoise biology and conservation.
However, LTDSMethod may be applicable for other studies/projects trying to
implement LTDS methods. 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(LTDSMethod)
```
