---
title: "LTDSMethod"
author: "Max Dolton Jones"
date: "`r Sys.Date()`"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{LTDSMethod}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Introduction to LTDSMethod
  
The _LTDSMethod_ package was designed for several main tasks:
  
* Designing line transect distance sampling pilot surveys
* Using pilot survey methodology to inform full LTDS surveys
* Estimate the density of a species within a study area
* Creating a summary of survey effort and results

More specifically, the package was initially intended as a tool to aid monitoring
efforts for translocated gopher tortoises throughout Florida. As a result, some
of the functions are highly specific to gopher tortoise biology and conservation.
However, _LTDSMethod_ may be applicable for other studies/projects trying to
implement LTDS methods. 

The _LTDSMethod_ package has a complimentary Standard Operating Procedure (SOP), 
which can be referenced to help understand background, terminology and access
some results behind survey design decisions. 

# Working example

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The _LTDSMethod_ package is currently open-access on a GitHub repository.
To download and install the package, we first need the devtools package.

```{r devtools, eval = F, echo = T}

#Install devtools: Issues? - https://github.com/r-lib/devtools/issues/2131
install.packages("devtools") #Ignore if previously installed
#Load devtools into the session:
library(devtools)

```

We can now use the devtools package to install _LTDSMethod_ from GitHub.

```{r github, eval = F, echo = T}

install_github("maxdoltonjones/LTDSMethod",  build_vignettes = TRUE, force = TRUE)

#The installation will ask if any package updates are needed, we can skip these
#by running the number three (3):
3

```

The installation of the package should also prompt the installation of all package dependencies (other packages which are needed for _LTDSMethod_ to work). However, if this is not the case, run the following lines (this may take a few minutes or so):

```{r installs, eval = F, echo = T}

install.packages("rgeos")
install("rgdal")
install("dplyr")
install.packages("readr")
install.packages("htmltools")
install.packages("gt")
install.packages("tinytex")
install.packages("nimble")
install.packages("terra")
install.packages("ggplot2")
install.packages("tinytex")
install.packages("tidyterra")
install.packages("knitr")
install.packages("imager")
install.packages("gstat")

```

Again, if these packages are already installed, the dependencies will be
loaded at the same time as _LTDSMethod_ is loaded. However, we can also load
all of the dependencies separately.

```{r, packages, eval = T, echo = T}

library(rgeos)
library(rgdal)
library(dplyr)
library(readr)
library(htmltools)
library(nimble)
library(terra)
library(ggplot2)
library(gt)
library(tinytex)
library(tidyterra)
library(knitr)
library(imager)
library(gstat)

```

The _LTDSMethod_ package works by saving and loading files from the working directory. It is important to set the working directory, and keep this consistent for the entirety of use of the package and functions. 

```{r wd, eval = F, echo = T}

setwd("FILEPATH HERE")

```

Now load in the _LTDSMethod_ package and set the session seed to follow the exact workflow of this vignette. If for some reason you need to run a function more than once, please re-run the set.seed function before re-running any functions. 

```{r setup, warning = FALSE}

library(LTDSMethod)

#Set a seed for reproducible results
set.seed(1235)

```

## Before sampling

Let's remind ourselves what is needed to work through the LTDSMethod package.
This information can be found in the Standard Operating Procedure (SOP):

1. Shapefile of recipient site (only permitted, habitat area)
1. Shapefile of additional acreage (only permitted, habitat area that is not
     already included in the original shapefile)
1. The acreage of the recipient site (only permitted, habitat area which is to
     be surveyed)
1. UTM zone of recipient site location (either 16 or 17, see Figure 3 of SOP)
1. Location of soft-release enclosures at recipient site (either embedded
     into shapefiles, or on a separate map)
     
Let's start with number 1. Read in our shapefile showing the recipient site to be surveyed.

```{r shp, eval = F, echo = T}

#We can read in a shapefile with:
site.shp <- readOGR(dsn = "SHAPEFILE FILE PATH HERE")
site.shp

```

If the data has been collected in FL Albers, or another projection system,
we can transform the data to work with the package via the next lines.
When setting the new.crs, make sure to allocate the appropriate zone using
the SOP as a guide. 

```{r, crs, eval = F, echo = T}

new.crs <- "+proj=utm +zone=17 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"
site.shp <- spTransform(site.shp, new.crs)
site.shp

#Convert the shapefile to a SpatVector, to work with the terra package
site.shp <- vect(site.shp)
#We can visually check the site by:
plot(site.shp)

```


We could move forward with the functions using just the site.shp shapefile just read in. However, here we are going to create our own shapefile and raster. We do this mostly for ease of reproducibility, however, it also allows us to easily visualize pen areas (5 on above list) which is useful for demonstrating the utility of the _LTDSMethod_ functions.
Specifically, we will create an irregular-shaped site (not square), with 100%
available permitted habitat area for tortoises - set at 400 acres, that also has two 120-acre soft-release pens.

```{r}

#400 acre irregular site, 100% habitat, 2x120 acre pens

#Site size is length in meters of one square edge
site.size <- 1500
#Pixel size will be set to 5m per pixel
pixel.size <- 5
#We can create a raster based on those set sizes. Projection is for Florida.
new.crs <- "+proj=utm +zone=17 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"
site <- rast(nrows=site.size/pixel.size, ncol=site.size/pixel.size, crs=new.crs)
#Set smallest Easting and Northing for site
xmin <- 300000
ymin <- 3350000
#Set the extent based on coordinates and size
ext(site) <- c(xmin, xmin+site.size, ymin, ymin+site.size)
#For now, set all pixel values to 1
values(site) <- 1
site

#The below code cuts out sections of the site to give an irregular shape.
area.1 <- crop(site, ext(300050, 301150, 3350050, 3350550))
area.2 <- crop(site, ext(300450, 301450, 3350550, 3351450))
area.3 <- crop(site, ext(300050, 300450, 3351050, 3351450))
site.new <- merge(area.1, area.2, area.3)

#We can use the below code to check the acreage of the site
area <- sum(values(site.new), na.rm = T)
(sqrt(area)*5)^2/4047

#We need to add our soft-release pens - just to help with the vignette example
pen.area <- site
values(pen.area) <- 3
pen.1 <- crop(pen.area, ext(300050, 301150, 3350050, 3350490))
pen.2 <- crop(pen.area, ext(300600, 301300, 3350650, 3351350))
pen.new <- merge(pen.1, pen.2)

#Merge site and pen area together
site.rast <- merge(pen.new, site.new)

#Convert and store shapefiles to crop rasters later
site.shp <- as.polygons(site.new)

#Plot raster to visualize site with pen areas
temp<-as.data.frame(site.rast, xy = T)
temp <- temp %>% na.omit()
ggplot(temp) +
  geom_tile(aes(x = x, y = y, fill = lyr.1)) +
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.8) +
  xlab("Easting") +
  ylab("Northing") +
  theme_classic() +
  theme(legend.position = "none")

```

We have now created a recipient site to work with. The second item on our list
is any additional acreage to be added to the site. We will only work with the 
site we have made above moving forward, but we can create additional acreage to
test package functionality later on. Additional acreage to the recipient site would usually be provided via shapefile, which we could load in as we did above. We are producing our own here for reproducibility purposes. 

```{r}
#Create another site, to demosntrate merging shapefiles
site.size <- 400  #Site size is length in meters of one square edge
pixel.size <- 5
new.crs <- "+proj=utm +zone=17 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"
site.2 <- rast(nrows=site.size/pixel.size, ncol=site.size/pixel.size, crs=new.crs)
xmin <- 301600
ymin <- 3350000
ext(site.2) <- c(xmin, xmin+site.size, ymin, ymin+site.size)
values(site.2) <- 1
site.2
#Convert and store shapefiles to crop rasters later
site.shp.2 <- as.polygons(site.2)

temp.2<-as.data.frame(site.2, xy = T)
temp.2 <- temp.2 %>% na.omit()
ggplot(temp.2) +
  geom_tile(aes(x = x, y = y, fill = lyr.1)) +
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.8) +
  xlab("Easting") +
  ylab("Northing") +
  theme_classic() +
  theme(legend.position = "none")

```

The last three requirements can be conveniently parsed from the first site we 
created. We have already checked the site acreage, but we do this again below.

```{r acreage}
#Calculate site acreage
area <- sum(values(site.new), na.rm = T)
(sqrt(area)*5)^2/4047

```

For number 4, most recipient sites in Florida will be in UTM zone 17, but
refer to Figure 3 of the SOP to check that your shapefile is not in zone 16. We will use zone 17 for this tutorial, which can be verified by:

```{r CRS}

#Call the shapefile object to check the site CRS (Coordinate Reference System)
site.rast

```

For number 5, we have created our own raster with pen edges, so we know where they are. A shapefile or raster file with pen locations would be the easiest method of visualizing pen locations. Otherwise, an aerial map or site visit would be another option. Plot raster as a reminder:

```{r pens}

ggplot(temp) +
  geom_tile(aes(x = x, y = y, fill = lyr.1)) +
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.8) +
  xlab("Easting") +
  ylab("Northing") +
  theme_classic() +
  theme(legend.position = "none")

```

If you have multiple shapefiles which need to be brought together to make
up the entire study site, we can do this using existing R functions. For
the below code to work, the shapefiles need to be converted into StatVector
objects using the vect() function. Then we can use rbind to merge them.

```{r, warning = FALSE}
#We can enter as many shapefiles as needed to merge into one
new_site <- rbind(site.shp, site.shp.2)
#We will use base plot to visualize this newly merged site
plot(new_site)
```

## Function:placing transect lines - _tran_place_ 

Check function help documentation:

```{r help2, eval=F, echo=T}

?tran_place

```

We are going to use our first function, _tran_place_, to delineate transect lines
across our recipient site depicting the lines to be walked for the pilot survey.
We need to provide the function with some key information, depending on the
size of the area to be surveyed (number 3 of checklist). We have a 400 acre site,
so we can refer to figure 4 of the SOP to know what characteristics are needed.
For this size, we need 50% of the overall transects to be surveyed for the pilot
survey, with a distance of 50m between adjacent transects.

Importantly, this is where we decide which direction the transect lines will be
running. The options are either North-South ("N-S"), East-West ("E-W"), Northeast-Southwest ("NE-SW"), or Northwest-Southeast ("NW-SE"). In an attempt to troubleshoot potential problems, we will run the function using a "N-S" orientation. We need to provide the site shapefile, UTM zone and if we want the lines to be plotted and saved to the working directory.

```{r}
#Set the distance between transects
dist <- 50

#The tran_place function produces a SpatVector object
pilot.lines <- tran_place(tran.dist = dist, pil.perc = 50, direction = "N-S",
                          site.poly = site.shp, zone = 17, plot = F, save = F)
#We can review information about the SpatVector object by running:
pilot.lines

#Plot the raster and lines to assess position relative to pens
temp<-as.data.frame(site.rast, xy = T)
temp <- temp %>% na.omit()
ggplot() +
  geom_tile(data = temp, aes(x = x, y = y, fill = lyr.1)) +
  geom_spatvector(data = pilot.lines, aes(),
               color = "black") +
  #geom_point(data = crd.test, aes(x = x, y = y)) +
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.8) +
  xlab("Easting") +
  ylab("Northing") +
  theme_classic() +
  theme(legend.position = "none")
```

We now have the first potential pilot lines that we could survey. However, when
choosing transect direction, it is extremely important to consider the position
of pen edges in relation to the transect lines. In our example, consider the 
following locations specifically:

```{r, echo = FALSE}

#Make points for data visualization
point.1 <- data.frame(
  x = c(300050, 300050, 300000, 300100),
  y = c(3350050, 3350500, 3350250, 3350250)
)
point.2 <- data.frame(
  x = c(300600, 300600, 300550, 300650),
  y = c(3350650, 3351300, 3351000, 3351000)
)
point.3 <- data.frame(
  x = c(301300, 301300, 301250, 301350),
  y = c(3350650, 3351300, 3351000, 3351000)
)
point.4 <- data.frame(
  x = c(301150, 301150, 301100, 301200),
  y = c(3350050, 3350500, 3350250, 3350250)
)
map.lines <- ggplot() +
  geom_tile(data = temp, aes(x = x, y = y, fill = lyr.1)) +
  geom_spatvector(data = pilot.lines, aes(),
                  color = "black") +
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.8) +
  xlab("Easting") +
  ylab("Northing") +
  theme_classic() +
  theme(legend.position = "none")

map.lines + stat_ellipse(data = point.1, aes(x = x, y = y), linetype = 2, level = 0.7,
                    color = "red") +
  stat_ellipse(data = point.2, aes(x = x, y = y), linetype = 2, level = 0.7,
               color = "red") +
  stat_ellipse(data = point.3, aes(x = x, y = y), linetype = 2, level = 0.7,
               color = "blue") +
  stat_ellipse(data = point.4, aes(x = x, y = y), linetype = 2, level = 0.7,
               color = "blue")
```

The dashed red lines show locations where the transect lines run directly on
top of pen edge locations. Where this occurs, we are likely to overestimate the
true density of tortoises. This is because translocated gopher tortoises are 
likely to cluster around these pen edges, because they typically offer built-up
sandy habitat (perfect for burrowing), but also limit outward movement - which
is typical for translocated reptiles. As a result of the transect line running
along one of these highly clustered edges, the encounter rate of a survey is 
inflated and ultimately the estimates are highly overestimated (SOP Figure 15). 

On the other hand, the blue dashed lines show where a pen edge has been
completely missed by the placement of transect lines. We have also found that
underestimations of density can occur due to an uneven sampling across pen edges, 
particularly where highly clustered pen edges are missed entirely (SOP Figure 15). 

Due to the poor placement of transect lines shown in the above figure, we will
now try running the _tran_place_ function again using a NE-SW direction. If we
tried an E-W direction, we could easily encounter the same issue as N-S, due to
the high linearity of our simulated study site. Let's see what the NE-SW looks
like:

```{r}
#Now trying the same parameters, but with NE-SW
pilot.lines <- tran_place(tran.dist = dist, pil.perc = 50, direction = "NE-SW",
                          site.poly = site.shp, zone = 17, plot = F, save = T)
#We can review information about the SpatialLines object by running:
pilot.lines

#Plot the raster and lines to assess position relative to pens
temp<-as.data.frame(site.rast, xy = T)
temp <- temp %>% na.omit()
ggplot() +
  geom_tile(data = temp, aes(x = x, y = y, fill = lyr.1)) +
  geom_spatvector(data = pilot.lines, aes(),
                  color = "black") +
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.8) +
  xlab("Easting") +
  ylab("Northing") +
  theme_classic() +
  theme(legend.position = "none")

```

These new pilot lines give a better distribution of transects across pen edges.
So we will be moving forward using these pilot transect lines.

## Function:getting start and end locations - _start_end_

Check function help documentation:

```{r help3, eval=F, echo=T}

?start_end

```

Now we have the transect lines needed for our pilot surveys, we are going to
extract the start and end locations of each line. These will be exported to a
GPS device so that we can adhere to line trajectories in the field.

For the start_end function to work, we need to provide the pilot lines
that we created using the _tran_place_ function, and the utm zone as before.
By setting save = T, this will save the points to the working directory.

```{r}
se.pts <- start_end(lines = pilot.lines, zone = 17, save = T, plot = F)

#Change to dataframe for plotting
se_loc <- data.frame(se.pts)
#We can also check the coordinated by plotting against the site and transects
ggplot() +
  geom_tile(data = temp, aes(x = x, y = y, fill = lyr.1)) +
  geom_spatvector(data = pilot.lines, aes(),
                  color = "black") +
  geom_point(data = se_loc, aes(x = Start_x, y = Start_y), size = 3, color = "orange") +
  geom_point(data = se_loc, aes(x = End_x, y = End_y), size = 3, color = "orange") +
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.8) +
  xlab("Easting") +
  ylab("Northing") +
  theme_classic() +
  theme(legend.position = "none")

```

The orange points now delineate either where a transect line should start or end
during a survey. Observers should survey from one orange point, to the opposite
orange point which is connected via the black transect line.

## Function:delineate locations for vegetation sampling - _veg_plot_

Check function help documentation:

```{r help4, eval=F, echo=T}

?veg_plot

```

While doing the surveys, collecting vegetation obstruction data can be
extremely important for accurate tortoise density estimation. In the next
function, we are going to determine where these vegetation plots are going
to be conducted.

The locations are determined in a very similar manner to the _start_end_ function
where we need to provide the pilot lines. However, we need to also provide
the set distance in meters between vegetation points (intv; default = 75m).
We provide the lines, via the start and end points created using the start_end function.

```{r}
#The csv file will be in the working directory
start.end <- read_csv("./start_end_coord.csv")
veg.loc <- veg_plot(points = start.end, intv = 75, save = T, plot = F)
#Change to dataframe for plotting
veg_loc <- data.frame(veg.loc)
#We can also check the coordinated by plotting against the site and transects
ggplot() +
  geom_tile(data = temp, aes(x = x, y = y, fill = lyr.1)) +
  geom_spatvector(data = pilot.lines, aes(),
                  color = "black") +
  geom_point(data = veg_loc, aes(x = coords.x1, y = coords.x2), size = 2,
             color = "red") +
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.8) +
  xlab("Easting") +
  ylab("Northing") +
  theme_classic() +
  theme(legend.position = "none")
```

The red points now show where observers will stop and collect vegetation
obstruction data. Consistent with our function, these are set at 75m apart. 

We have come to the end of our "before sampling" section. Let's review our working directory folders to see if we have everything needed to get out in the field and perform our pilot survey.

Below shows all the files we have in our working directory. We need the following files to continue:

1. Transect_lines.shp
1. Start_end_locations.shp
1. Veg_plots.shp

```{r list, eval=F, echo=T}

#List all files in working directory
list.files()

```

## After pilot

## Function:calculate pilot effort - _samp_effort_

Check function help documentation:

```{r help5, eval=F, echo=T}

?samp_effort

```

Once the pilot survey has been conducted, as outlined in the SOP, we can take
the start and end locations for all surveyed lines and calculate the effort.
Our _start_end_ function creates a csv file with start and end locations for
each transect line. This mirrors the expected dataframe from an actual field
survey. The order of the columns is the most important for the function.
See the SOP for information on how to properly format the coordinate data
during and following an LTDS survey.

To calculate the effort (distance in meters), we provide the start and end
locations of each transect line walked during the pilot survey. As before, we
also need the utm zone. We also have the option to plot the transect lines.
Since we are using lines that we created, we will not re-plot the lines, however, 
this could give us the opportunity to assess if there are any errors
with the data collection or function. For example, if the lines do not look
anything like what was actually walked (besides being perfectly straight), then
there is likely an issue.

```{r}
#The csv file will be in the working directory
start.end <- read_csv("./start_end_coord.csv")

effort <- samp_effort(points = start.end, zone = 17, plot = F)

#View resulting effort - this is another place to check if the distance is
#corresponding to the actual effort in the field.
effort
```

As we can see from above, the effort for the pilot survey was 19464m (~19.5km).

## Function:determine lines for full LTDS survey - _samp_full_

Check function help documentation:

```{r help6, eval=F, echo=T}

?samp_full

```

Now we have the distance covered during the pilot survey, we can calculate the
remaining effort needed to attain our target CV (default = 17%). The following
function works similarly to the _tran_place_ function used at the
beginning of the tutorial. However, we do not want to duplicate effort, so we
now build upon the pilot survey to add additional transect lines. This means that the pilot line effort will be included within the final density estimation. 

There are several bits of information that we need to provide to get the function
to work. Firstly, we need the effort from the _samp_effort_ function. We need to
state the direction of the transects (which needs to match the direction of the
pilot survey), which in this case is NE-SW. Next, we need the number of occupied
burrows detected during the pilot survey. If there were a lot of occupied
burrows discovered, we can pull this information from the data. However, for the
purpose of this vignette, we will set the number of occupied burrows (ntort) to 
120. We then need to provide the desired CV from the survey. For our tutorial, we
will aim for a CV of 17%. As with the other functions, we lastly need to provide
the utm zone.

```{r}
full.ltds <- samp_full(effort = effort, site.poly = site.shp, direction = "NE-SW",
                       ntort = 120, cv = 0.17, zone = 17, plot = F, save = F)
```

By setting the ntort to 120, the object is not created. In the console,
we can see the statement "Target CV attained". This means that by finding 120 occupied burrows with the current amount of pilot effort, we have already achieved a CV
that is lower than our target. In this case, the CV is 15.81%. Now, let's try the opposite, what if we only detected 50 occupied burrows?

```{r}
full.ltds <- samp_full(effort = effort, site.poly = site.shp, direction = "NE-SW",
                       ntort = 50, cv = 0.17, zone = 17, plot = F, save = F)

ggplot() +
  geom_tile(data = temp, aes(x = x, y = y, fill = lyr.1)) +
  geom_spatvector(data = pilot.lines, aes(),
                  color = "black") +
  geom_spatvector(data = full.ltds, aes(),
                  color = "red") +
  
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.8) +
  xlab("Easting") +
  ylab("Northing") +
  theme_classic() +
  theme(legend.position = "none")
```

With this much lower encounter rate, our CV was much higher at 24.49%. We can see
from the accompanying plot that we would now need to survey all possible
transects to achieve a more adequate CV.

Let's set the number of detected burrows to 100 to move forward with the vignette example.

```{r}

full.ltds <- samp_full(effort = effort, site.poly = site.shp, direction = "NE-SW",
                       ntort = 100, cv = 0.17, zone = 17, plot = F, save = F)

ggplot() +
  geom_tile(data = temp, aes(x = x, y = y, fill = lyr.1)) +
  geom_spatvector(data = pilot.lines, aes(),
                  color = "black") +
  geom_spatvector(data = full.ltds, aes(),
                  color = "red") +
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.8) +
  xlab("Easting") +
  ylab("Northing") +
  theme_classic() +
  theme(legend.position = "none")

```

By assigning 100 discovered, occupied burrows we reached a CV of 17.32%. The 
function has delineated new lines (red), that need to be surveyed to attain a CV
under 17%. We will move forward in the vignette using the lines from discovering 100 burrows. 

As with the pilot transects, we need to know the start and end locations to make it easier to find and traverse the transect lines in the field. To do this, we can simply re-run the _start_end_ function again using our full.ltds lines.

_IMPORTANT_ - Before running the next lines, rename the start_end files so that they are not overwritten by the function.

```{r}
se.pts <- start_end(lines = full.ltds, zone = 17, save = T, plot = F)

#We can also check the coordinates by plotting against the site and transects
#Change to dataframe for plotting
se_loc <- data.frame(se.pts)
#We can also check the coordinated by plotting against the site and transects
ggplot() +
  geom_tile(data = temp, aes(x = x, y = y, fill = lyr.1)) +
  geom_spatvector(data = pilot.lines, aes(),
                  color = "black") +
  geom_spatvector(data = full.ltds, aes(),
                  color = "red") +
  geom_point(data = se_loc, aes(x = Start_x, y = Start_y), size = 3, color = "orange") +
  geom_point(data = se_loc, aes(x = End_x, y = End_y), size = 3, color = "orange") +
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.8) +
  xlab("Easting") +
  ylab("Northing") +
  theme_classic() +
  theme(legend.position = "none")

```

Lastly, we need to know the location for collecting vegetation obstruction data as before. Again, re-run the veg_plot function using the full.ltds lines to delineate the locations.

_IMPORTANT_ - Before running the next lines, rename the veg_plot files so that they are not overwritten by the function.

```{r}
#Re-run the veg_plot function, using the new start_end csv in the wd.
start.end <- read_csv("./start_end_coord.csv")
veg.loc <- veg_plot(start.end, intv = 75, save = T, plot = F)

#Change to dataframe for plotting
veg_loc <- data.frame(veg.loc)
ggplot() +
  geom_tile(data = temp, aes(x = x, y = y, fill = lyr.1)) +
  geom_spatvector(data = full.ltds, aes(),
                  color = "black") +
  geom_point(data = veg_loc, aes(x = coords.x1, y = coords.x2), size = 2,
             color = "red") +
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.8) +
  xlab("Easting") +
  ylab("Northing") +
  theme_classic() +
  theme(legend.position = "none")
```

We have come to the end of our "after pilot" section. Let's review our working directory folders to see if we have everything needed to get out in the field and perform our full LTDS survey.

Below shows all the files we have in our working directory. We need the following files to continue:

1. Add_transect_lines.shp
1. Start_end_locations.shp
1. Veg_plots.shp

```{r list2, eval=F, echo=T}

#List all files in working directory
list.files()

```

## After full LTDS

We now presume that the full LTDS survey has been conducted. This survey will
ultimately give us a dataframe with several important variables. Import the
data by:

```{r}
full.data <- read_csv("./LTDS_example_data_new.csv")

#We can summarize the data
summary(full.data)
```

## Function:truncate data - _LTDS_crop_

Check function help documentation:

```{r help7, eval=F, echo=T}

?LTDS_crop

```

So that we can use the data to estimate the density of tortoises at the site,
we are going to slightly format the input. Specifically, we are going to
truncate the data so that any burrows found over half the distance between the
lines are removed; and we are also going to remove the top distance outliers (5%).

The function we are using simply needs two bits of information. We need the
data from the full survey, and we also need to set the distance between transects, which will determine the truncation distance. We have created the object "dist" earlier in the vignette which we can input into the function.

```{r}
data.new <- LTDS_crop(ltds_data = full.data, tran.dist = dist)

#We can summarize the data again for comparison.
summary(data.new)
```

## Function:estimate the density - _dens_est_

Check function help documentation:

```{r help8, eval=F, echo=T}

?dens_est

```

We are finally at the penultimate function of the _LTDSMethod_ package. We are
now ready to calculate the density of tortoises at the recipient site.

For the function to work and estimate the density, we need to provide three
main bits of information. We need the cropped ltds data from _LTDS_crop_, the
effort resulting from _samp_effort_, and lastly we need the total area of the
recipient site in acres.

Remember, we have now conducted the full LTDS survey, which means that our resulting effort will be greater than the _samp_effort_ function has previously calculated (unless no further transect lines were walked during the full LTDS; i.e., the target CV had been attained during the pilot). We therefore need to re-run the _samp_effort_ function using the additional lines. You will need to combine the start and end csv files from the pilot and full survey to calculate the correct amount of effort. 

For this vignette, we have provided a combined start and end coordinate csv file to work with. This can be read in as before:

```{r}
#The csv file will be in the working directory
start.end.2 <- read_csv("./start_end_coord_full.csv")

effort.2 <- samp_effort(points = start.end.2, zone = 17, plot = F)

#View resulting effort - this is another place to check if the distance is
#corresponding to the actual effort in the field.
effort.2
```

The final effort for the full LTDS survey was 21299m (~21.3km). We can enter the effort.2 object into our function.

```{r}
#Calculate density
tort.dens <- dens_est(ltds_data = data.new, effort = effort.2, area = 400)

#The dens_est function produces a small dataframe with the estimated density
#and 95% credible interval around that estimate. This gives us a range of 
#possible densities at the recipient site. 
tort.dens$density #Mean estimate
tort.dens$density.lo #95% CI Low
tort.dens$density.hi #95% CI High

#This means that our total number of tortoises estimated at the site is:
tort.pop <- tort.dens*400 #Change number according to acreage at site
tort.pop$density #Mean estimate
tort.pop$density.lo #95% CI Low
tort.pop$density.hi #95% CI High

```

Our final density estimate at the site is 1.57 tortoises per acre with a range of 1.28 - 1.92 tortoises/acre based on our 95% credible interval values. This means that there is an estimated approximately 629 tortoises at our site, with a possible range of 513-767 tortoises. This is fabricated data to fit in with the vignette training, so the accuracy of this estimation is unknown. See the SOP for simulation results.

What if there was an obstruction on one of the transect lines? Instructions for encountering an obstruction in the field (such as a flooded area or fallen tree), can be found in the accompanying SOP. The new coordinates will be read in as before:

```{r}
start.end.3 <- read_csv("./start_end_coord_obst.csv")

```

These depict the following effort in the field:

```{r, echo = FALSE}

start.end.3 <- read_csv("./start_end_coord_obst.csv")
l <- vector("list", nrow(start.end.3))
for(i in 1:nrow(start.end.3)){
  l[[i]] <- Lines(list(Line(cbind(c(as.numeric(start.end.3[i, 1]), as.numeric(start.end.3[i, 3])),
                                  c(as.numeric(start.end.3[i, 2]), as.numeric(start.end.3[i, 4]))))),
                  as.character(i))
}
lines <- SpatialLines(l)
#zone <- 17
crs(lines) <- crs(paste("+proj=utm + zone=", 17, " ellps=WGS84", sep=''))

point.1 <- data.frame(
  x = c(301050, 301050, 301200, 301200),
  y = c(3350630, 3350500, 3350600, 3350880)
)

data <- data.frame(lines=1:length(lines))
obst.lines <- SpatialLinesDataFrame(lines, data, match.ID = FALSE)
obst.plot <- ggplot() +
  geom_tile(data = temp, aes(x = x, y = y, fill = lyr.1)) +
  geom_polygon(data = obst.lines, aes(x = long, y = lat, group = group),
               color = "black") +
  scale_fill_viridis_c(option = "mako", begin = 0.3, end = 0.8) +
  xlab("Easting") +
  ylab("Northing") +
  theme_classic() +
  theme(legend.position = "none")

obst.plot + stat_ellipse(data = point.1, aes(x = x, y = y), linetype = 2, level = 0.6,
                    color = "red", size = 1)

```

Our new effort can be calculated as before.

```{r}
#The csv file will be in the working directory
start.end.3 <- read_csv("./start_end_coord_obst.csv")

effort.3 <- samp_effort(points = start.end.3, zone = 17, plot = F)

#View resulting effort - this is another place to check if the distance is
#corresponding to the actual effort in the field.
effort.3
```

The effort when factoring in the site obstruction is now 21101m (21.1km). Which can be included in the _dens_est_ function as above. 

## Function:produce LTDS final report - _LTDS_summary_

Now we can produce our final report. There is slightly more manual input for
the report, since we wanted to make this adaptable to changes in data and
results which cannot be captured by the LTDSMethod package. The parameters of
the function and report are as follows:

To move forward, we need to read in our other example dataset. During the
surveys, observers have the opportunity to record an other observations
made. Specifically, any observations of commensal species, or discoveries
of gopher tortoise carcasses. 

```{r}
obs.data <- read_csv("./Other_observation_form_v2.csv")
#Summarize the data
head(obs.data)
#effort: the effort from the full LTDS survey. We will use the pilot effort that
#we calculated earlier for this example.
effort <- as.numeric(effort.2)
#nburr: the total number of burrows found during the survey. We can axtract this
#from our example dataset, which will likely be applicable for future use.
nburr <- as.numeric(length(full.data$Burrow_width))
#burr.w.mn: the mean width of detected burrows.
burr.w.mn <- as.numeric(round(mean(full.data$Burrow_width), digits = 2))
#burr.w.min: the minimum width of detected burrows.
burr.w.min <- as.numeric(round(min(full.data$Burrow_width), digits = 2))
#burr.w.max: the maximum width of detected burrows.
burr.w.max <- as.numeric(round(max(full.data$Burrow_width), digits = 2))
#comm.comm: the name of the most common commensal species detected. We can
#extract the number of detections for the full dataset by:
table(obs.data$Species)
#This shows that the most common species was the Florida mouse
comm.comm <- "Florida mouse"
#occ: the number of occupied burrows discovered.
#unocc: the number of unoccupied burrows discovered.
#unk: the number of unknown burrows discovered.
#We can use the same method as above to pull these numbers out of the data
table(full.data$Occupied)
occ <- 107
unocc <- 131
unk <- 25
#site.name: the name of the recipient site. We can set this to whatever we want.
site.name <- "Tortoise Wonderland"
#site.size: total area of site, or sampling frame (acres)
site.size <- 400
#tort.dens: the estimated tortoise density at the site. We calculated this earlier.
tort.dens.new <- round(tort.dens$density, digits = 2)
tort.dens.lo <- round(tort.dens$density.lo, digits = 2)
tort.dens.hi <- round(tort.dens$density.hi, digits = 2)
#tort.pop: the estimated tortoise population at the site. We calculated this earlier.
tort.pop.new <- round(tort.pop$density)
tort.pop.lo <- round(tort.pop$density.lo)
tort.pop.hi <- round(tort.pop$density.hi)

```

The above information is used for the text part of the documents, however, any
figures produced in the report will pull data directly from the working directory.

We can now input all of these into our function. As a warning, if you have
previously run this function, and viewed the resulting pdf, the function will
not run if you still have the pdf open.

```{r eval=FALSE}
LTDS_summary(effort = effort, nburr = nburr, burr.w.mn = burr.w.mn,
             burr.w.min = burr.w.min, burr.w.max = burr.w.max, comm.comm = comm.comm,
             occ = occ, unocc = unocc, unk = unk, site.name = site.name,
             site.size = site.size,
             tort.dens = tort.dens.new, tort.dens.lo = tort.dens.lo,
             tort.dens.hi = tort.dens.hi, tort.pop = tort.pop.new,
             tort.pop.lo = tort.pop.lo, tort.pop.hi = tort.pop.hi)

```

End of vignette.

